Manuscript

### 2018-12-05 manuscript.Rmd
start on lineage-specific analysis for manuscript

```{r}
WilcoxonByDrug <- function(drug, dataset, data_name, grouping_var) {
  if(grouping_var == "lineage") {
    grouper <- c("Hugo_Symbol", "group_general_lineage_name")
  }
  else if(grouping_var == "gene") {
    grouper <- "Hugo_Symbol"
  }
  use_data <- filter(dataset, Drug == drug)
  sig <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = grouper, data = use_data, method = "wilcox.test", p.adjust.method = "BH")
  sig <- adj_signif(sig)
  sig <- sig[order(sig$p),]
  sig$Drug <- drug
  sig$Dataset <- toupper(data_name)
  return(sig)
}
# WilcoxonByDrugAllGenes <- function(melt_file, grid_file, data_name) {
#   druglist <- as.character(unique(melt_file$Drug))
#   sig_df <- NULL
#   for(i in 1:length(druglist)) {
#     print(paste0("Current drug in process: ", druglist[i], " (drug ", i, " of ", length(druglist), ")."))
#     flush.console()
#     melted <- filter(melt_file, Drug == druglist[i])
#     use_data <- merge(grid_file, melted, by = "accession_id", all.x = TRUE)
#     use_data <- merge(use_data, ccl_converter, by = "accession_id", all.x = TRUE)
#     use_data <- merge(use_data, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
#     use_data$Mutation_Status_Nonsilent <- ifelse(is.na(use_data$Mutation_Status_Nonsilent), "Wildtype", use_data$Mutation_Status_Nonsilent)
#     use_data$Hugo_Symbol <- factor(use_data$Hugo_Symbol)
#     use_data$AUC <- as.numeric(use_data$AUC)
#     use_data <- filter(use_data, !is.na(use_data$AUC))
#     use_data$Drug_Gene <- paste0(use_data$Drug, "_", use_data$Hugo_Symbol)
#     use_data_ptmuts <- filter(use_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
#     use_data_ptmuts$Keep_PtMut_Tests <- ifelse(use_data_ptmuts$Reference_Allele_Length == 1 | is.na(use_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(use_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
#     use_data_ptmuts <- filter(use_data_ptmuts, Keep_PtMut_Tests == TRUE)
#     
#     sig <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Hugo_Symbol", data = use_data_ptmuts, method = "wilcox.test", p.adjust.method = "BH")
#     sig <- adj_signif(sig)
#     sig <- sig[order(sig$p),]
#     sig <- filter(sig, p <= 0.05)
#     sig$Drug <- druglist[i]
#     sig$Dataset <- toupper(data_name)
#     sig_df <- rbind(sig_df, sig)
#   }
#   return(sig_df)
# }
```

@@

```{r}
g2p_nonCID <- subset(g2p, !grepl("^CID", g2p$DrugID))
summ <- g2p_nonCID %>% group_by(Drug, DrugID) %>% tally() %>% summarize(DrugID = DrugID, n = n, percent = formatC(n / 402 * 100, digits = 6, format = "f"))
summ
```

@@

@@

@@

crispr_data_signif_g2p_lineage_plot <- ggplot(data = crispr_data_signif_g2p_lineage, aes(x = group_general_lineage_name, y = Score)) +
  facet_wrap(~ Hugo_Symbol, nrow = 6, scales = "free_y") +
  geom_point(aes(color = Mutation_Status_Nonsilent), alpha = 0.5) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = c("Mutant" = "darkorchid", "Wildtype" = "thistle")) +
  geom_text(data = crispr_data_signif_g2p_lineage_text, mapping = aes(x = group_general_lineage_name, y = Text_y, label = p.signif), angle = 90, vjust = 0.8, nudge_y = 0.1) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES score by cell line lineage for significant Level A G2P genes",
       subtitle = "* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\nGenes are sorted by increasing lineage-agnostic p-value.")
crispr_data_signif_g2p_lineage_plot

@@

```{r}
ccle_signif_g2p_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")
knitr::kable(ccle_signif_g2p_lineage[1:500, c("Hugo_Symbol", "Drug", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: Wilcoxon test results for level A G2P drug-gene associations across lineages (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

@@

g2p_gene_tally <- g2p_filt %>% group_by(Gene) %>% tally() %>% drop_na()
range(g2p_gene_tally$n)
mean(g2p_gene_tally$n)
sd(g2p_gene_tally$n)
```

@@

### All

@@

Pearson summary stats:
```{r fig.height=4, fig.width=5}
g2p_druggene_pearson_stats <- g2p_druggene_pearson
g2p_druggene_pearson_stats$n_true <- TRUE
g2p_druggene_pearson_stats <- g2p_druggene_pearson_stats %>% group_by(Dataset, InG2P) %>% drop_na() %>% summarize(
          ge_mean = round(mean(Pearson_GE_Abs), 5),
          ge_sd = round(sd(Pearson_GE_Abs), 5),
          ge_n = sum(n_true),
          ge_n_signif = sum(Pearson_Pval_GE < 0.05),
          ge_n_sig_perc = ge_n_signif / ge_n * 100,
          cn_mean = round(mean(Pearson_CN_Abs), 3),
          cn_sd = round(sd(Pearson_CN_Abs), 5),
          cn_n = sum(n_true),
          cn_n_signif = sum(Pearson_Pval_CN < 0.05),
          cn_n_sig_perc = cn_n_signif / cn_n * 100)

ctrp_grid_summary_data <- filter(ctrp_data_g2p_grid, Drug_Gene == "afatinib_ERBB2")

@@

>>>>

```{r}
WilcoxonByDrug <- function(drug, dataset, data_name, grouping_var) {
  if(grouping_var == "lineage") {
    grouper <- c("Hugo_Symbol", "group_general_lineage_name")
  }
  else if(grouping_var == "gene") {
    grouper <- "Hugo_Symbol"
  }
  use_data <- filter(dataset, Drug == drug)
  sig <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = grouper, data = use_data, method = "wilcox.test", p.adjust.method = "BH")
  sig <- adj_signif(sig)
  sig <- sig[order(sig$p),]
  sig$Drug <- drug
  sig$Dataset <- toupper(data_name)
  return(sig)
}

# WilcoxonByDrugAllGenes <- function(melt_file, grid_file, data_name) {
#   druglist <- as.character(unique(melt_file$Drug))
#   sig_df <- NULL
#   for(i in 1:length(druglist)) {
#     print(paste0("Current drug in process: ", druglist[i], " (drug ", i, " of ", length(druglist), ")."))
#     flush.console()
#     melted <- filter(melt_file, Drug == druglist[i])
#     use_data <- merge(grid_file, melted, by = "accession_id", all.x = TRUE)
#     use_data <- merge(use_data, ccl_converter, by = "accession_id", all.x = TRUE)
#     use_data <- merge(use_data, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
#     use_data$Mutation_Status_Nonsilent <- ifelse(is.na(use_data$Mutation_Status_Nonsilent), "Wildtype", use_data$Mutation_Status_Nonsilent)
#     use_data$Hugo_Symbol <- factor(use_data$Hugo_Symbol)
#     use_data$AUC <- as.numeric(use_data$AUC)
#     use_data <- filter(use_data, !is.na(use_data$AUC))
#     use_data$Drug_Gene <- paste0(use_data$Drug, "_", use_data$Hugo_Symbol)
#     use_data_ptmuts <- filter(use_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
#     use_data_ptmuts$Keep_PtMut_Tests <- ifelse(use_data_ptmuts$Reference_Allele_Length == 1 | is.na(use_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(use_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
#     use_data_ptmuts <- filter(use_data_ptmuts, Keep_PtMut_Tests == TRUE)
#     
#     sig <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Hugo_Symbol", data = use_data_ptmuts, method = "wilcox.test", p.adjust.method = "BH")
#     sig <- adj_signif(sig)
#     sig <- sig[order(sig$p),]
#     sig <- filter(sig, p <= 0.05)
#     sig$Drug <- druglist[i]
#     sig$Dataset <- toupper(data_name)
#     sig_df <- rbind(sig_df, sig)
#   }
#   return(sig_df)
# }
```

@@

```{r}
# levels(g2p_filt$Phenotype.Description)
unique(g2p_filt$group_general_lineage_name)
levels(g2p_filt$group_general_lineage_name)

g2p_nonCID <- subset(g2p, !grepl("^CID", g2p$DrugID))
summ <- g2p_nonCID %>% group_by(Drug, DrugID) %>% tally() %>% summarize(DrugID = DrugID, n = n, percent = formatC(n / 402 * 100, digits = 6, format = "f"))
summ
```

@@

```{r}
length(unique(ccle_data_g2p_grid$Drug))
length(unique(ccle_data_g2p_grid$Drug_Gene))
```

@@

```{r}
unique(crispr_data_g2p$lineage_name)
```

@@

crispr_data_signif_g2p_lineage_plot <- ggplot(data = crispr_data_signif_g2p_lineage, aes(x = group_general_lineage_name, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 6, scales = "free_y") +
  geom_point(aes(color = Mutation_Status_Nonsilent), alpha = 0.5) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  scale_color_manual(values = c("Mutant" = "darkorchid", "Wildtype" = "thistle")) +
  geom_text(data = crispr_data_signif_g2p_lineage_text, mapping = aes(x = group_general_lineage_name, y = Text_y, label = p.signif), angle = 90, vjust = 0.8, nudge_y = 0.1) +
  theme(legend.position = "top", axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4, size = 10)) +
  labs(y = "CERES Score",
       x = "Lineage",
       color = "Mutation Status",
       title = "CERES score by cell line lineage for significant Level A G2P genes",
       subtitle = "* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001\nGenes are sorted by increasing lineage-agnostic p-value.")
crispr_data_signif_g2p_lineage_plot

@@

```{r}
ccle_signif_g2p_lineage <- readRDS("./data_munging/rds/ccle_signif_g2p_lineage.rds")
knitr::kable(ccle_signif_g2p_lineage[1:500, c("Hugo_Symbol", "Drug", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: Wilcoxon test results for level A G2P drug-gene associations across lineages (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")

knitr::kable(ccle_signif_g2p_lineage[1:500, c("Hugo_Symbol", "Drug", "group_general_lineage_name", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: Wilcoxon test results for level A G2P drug-gene associations across lineages (* p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

@@

g2p_gene_tally <- g2p_filt %>% group_by(Gene) %>% tally() %>% drop_na()
range(g2p_gene_tally$n)
mean(g2p_gene_tally$n)
sd(g2p_gene_tally$n)

g2p_druggenelin_tally <- g2p_filt %>% group_by(Drug, Gene, Drug_Gene, group_general_lineage_name) %>% tally() %>% drop_na()
length(unique(g2p_druggenelin_tally$Drug))
length(unique(g2p_druggenelin_tally$Gene))
unique(g2p_filt$Phenotype.Description)
```

@@

### Mut, GE, and CN

#### Lineage-agnostic

@@

Pearson summary stats:
```{r fig.height=4, fig.width=5}
g2p_druggene_pearson_stats <- g2p_druggene_pearson
g2p_druggene_pearson_stats$n_true <- TRUE
g2p_druggene_pearson_stats <- g2p_druggene_pearson_stats %>% group_by(Dataset, InG2P) %>% drop_na() %>% summarize(
          ge_mean = round(mean(Pearson_GE_Abs), 5),
          ge_sd = round(sd(Pearson_GE_Abs), 5),
          ge_n = sum(n_true),
          ge_n_signif = sum(Pearson_Pval_GE < 0.05),
          ge_n_sig_perc = ge_n_signif / ge_n * 100,
          cn_mean = round(mean(Pearson_CN_Abs), 3),
          cn_sd = round(sd(Pearson_CN_Abs), 5),
          cn_n = sum(n_true),
          cn_n_signif = sum(Pearson_Pval_CN < 0.05),
          cn_n_sig_perc = cn_n_signif / cn_n * 100)
```

Example p-value and Pearson correlation coefficient derivations:
```{r}
ctrp_grid_summary_data <- filter(ctrp_data_g2p_grid, Drug_Gene == "afatinib_ERBB2")

@@

#### By lineage

Deriving n for Bonferroni correction:
```{r}
ttest_summ_res_mut <- g2p_druggenelin_p_ttest_res_list[, c("p_value", "Dataset", "group_general_lineage_name")]
ttest_summ_res_mut$Metric <- "Mutation Status"
mut_alphas <- ttest_summ_res_mut %>% drop_na(p_value) %>% group_by(Dataset) %>% tally()
mut_alphas$alpha_adj <- 0.05 / mut_alphas$n

ttest_summ_res_ge <- druggenelin_pearson_ttest_res_list[, c("p_value_GE", "Dataset", "group_general_lineage_name")]
colnames(ttest_summ_res_ge) <- c("p_value", "Dataset", "group_general_lineage_name")
ttest_summ_res_ge$Metric <- "Gene Expression"
ge_alphas <- ttest_summ_res_ge %>% drop_na(p_value) %>% group_by(Dataset) %>% tally()
ge_alphas$alpha_adj <- 0.05 / ge_alphas$n

ttest_summ_res_cn <- druggenelin_pearson_ttest_res_list[, c("p_value_CN", "Dataset", "group_general_lineage_name")]
colnames(ttest_summ_res_cn) <- c("p_value", "Dataset", "group_general_lineage_name")
ttest_summ_res_cn$Metric <- "Copy Number"
cn_alphas <- ttest_summ_res_cn %>% drop_na(p_value) %>% group_by(Dataset) %>% tally()
cn_alphas$alpha_adj <- 0.05 / cn_alphas$n
```

Facted lineage summaries:
```{r fig.height=4, fig.width=12}
ttest_mut_summ_facet <- ggplot(data = ttest_summ_res, aes(x = p_value, fill = Dataset, color = Dataset)) +
  facet_grid(Metric ~ Dataset) +
  geom_histogram(bins = 40, binwidth = 0.025, boundary = 0, closed = "left") +
  scale_y_continuous(breaks = c(0:6), labels = c(0:6)) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.1), expand = c(0.01, 0.01)) +
  scale_fill_manual(values = c("CCLE" = "palegreen3", "CTRP" = "slategray3", "GDSC" = "paleturquoise3")) +
  scale_color_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  theme(legend.position = c(0.5, 1.7), legend.direction = "horizontal", panel.grid.minor.y = element_blank()) +
  labs(x = "P-value", y = "Frquency")
ttest_mut_summ_facet

# ggsave("./plots/manuscript/all_druggenelin_ttest_summ_facet_plot.png", ttest_mut_summ_facet, device = "png", dpi = 450, width = 12, height = 4, units = "in")
```

Stacked lineage summaries:
```{r fig.height=7, fig.width=12}
# margin(t = 0, r = 0, b = 0, l = 0, unit = "pt")
ttest_summ_res <- rbind(ttest_summ_res_mut, ttest_summ_res_ge, ttest_summ_res_cn)
ttest_summ_res$Metric <- factor(ttest_summ_res$Metric, levels = c("Mutation Status", "Gene Expression", "Copy Number"))

ttest_mut_summ_plot <- ggplot(data = g2p_druggenelin_p_ttest_res_list, aes(x = p_value)) +
  geom_histogram(aes(fill = Dataset, color = Dataset), position = "stack", bins = 40, binwidth = 0.025, boundary = 0, closed = "left") +
  scale_y_continuous(breaks = c(0:6), labels = c(0:6)) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), expand = c(0.01, 0.01)) +
  scale_fill_manual(values = c("CCLE" = "palegreen3", "CTRP" = "slategray3", "GDSC" = "paleturquoise3")) +
  scale_color_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  theme(legend.position = c(0.5, 1.7), legend.direction = "horizontal", panel.grid.minor.y = element_blank(), panel.grid.minor.x = element_blank(), plot.margin = unit(c(30, 5.5, 1, 5.5), "pt")) +
  labs(x = "P-value", y = "Frquency", title = "Mutation status")

ttest_ge_summ_plot <- ggplot(data = druggenelin_pearson_ttest_res_list, aes(x = p_value_GE)) +
  geom_histogram(aes(fill = Dataset, color = Dataset), position = "stack", bins = 40, binwidth = 0.025, boundary = 0, closed = "left") +
  scale_y_continuous(breaks = c(0:7), labels = c(0:7), expand = c(0.01, 0.01)) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), expand = c(0.01, 0.01)) +
  scale_fill_manual(values = c("CCLE" = "palegreen3", "CTRP" = "slategray3", "GDSC" = "paleturquoise3")) +
  scale_color_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  theme(legend.position = "none", panel.grid.minor.y = element_blank(), panel.grid.minor.x = element_blank(), plot.margin = unit(c(1, 5.5, 1, 5.5), "pt")) +
  labs(x = "P-value", y = "Frquency", title = "Gene expression")

ttest_cn_summ_plot <- ggplot(data = druggenelin_pearson_ttest_res_list, aes(x = p_value_CN)) +
  geom_histogram(aes(fill = Dataset, color = Dataset), position = "stack", bins = 40, binwidth = 0.025, boundary = 0, closed = "left") +
  scale_y_continuous(breaks = c(0:6), labels = c(0:6), expand = c(0.01, 0.01)) +
  scale_x_continuous(breaks = seq(0, 1, by = 0.05), expand = c(0.01, 0.01)) +
  scale_fill_manual(values = c("CCLE" = "palegreen3", "CTRP" = "slategray3", "GDSC" = "paleturquoise3")) +
  scale_color_manual(values = c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")) +
  theme(legend.position = "none", panel.grid.minor.y = element_blank(), panel.grid.minor.x = element_blank(), plot.margin = unit(c(1, 5.5, 5.5, 5.5), "pt")) +
  labs(x = "P-value", y = "Frquency", title = "Copy number")

ttest_mut_summ_plot <- ggplot_gtable(ggplot_build(ttest_mut_summ_plot))
ttest_ge_summ_plot <- ggplot_gtable(ggplot_build(ttest_ge_summ_plot))
ttest_cn_summ_plot <- ggplot_gtable(ggplot_build(ttest_cn_summ_plot))

# p_maxWidth <-  grid::unit.pmax(ttest_mut_summ_plot$widths, ttest_ge_summ_plot$widths, ttest_cn_summ_plot$widths)
# ttest_mut_summ_plot$widths <- as.list(p_maxWidth)
# ttest_ge_summ_plot$widths <- as.list(p_maxWidth)
# ttest_cn_summ_plot$widths <- as.list(p_maxWidth)

ttest_summ_plot <- cowplot::plot_grid(ttest_mut_summ_plot, ttest_ge_summ_plot,ttest_cn_summ_plot, nrow = 3, rel_heights = c(1, 1.6, 1.6))
ttest_summ_plot

# ggsave("./plots/manuscript/all_druggenelin_ttest_summ_plot.png", ttest_summ_plot, device = "png", dpi = 450, width = 12, height = 7, units = "in")
```
